---
title: "Attach_Phenodata"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
library(SummarizedExperiment)
library(readxl)
library(dplyr)
```
```{r}
load('RTI/rse_gene_Jlab_experiment_n242.Rdata')
rse_md=as.data.frame(colData(rse_gene))

# read the metadata from an Excel file:
phenodata <- read_xlsx("RTI/RTI2021.xlsx")
```

```{r}
rse_md$RNum <- gsub('^(R\\d+).*', '\\1', rse_md$SAMPLE_ID, perl=T)

cols <- c('RNum', 'RIN', 'BrNum', 'Brain Region', 'PrimaryDx', 'Sex', 'Race', 'AgeDeath' )
rncols <- c('RNum', 'RIN', 'BrNum', 'Region', 'Dx', 'Sex', 'Race', 'Age' )
pdadd <- dplyr::select(phenodata, all_of(cols))
colnames(pdadd) <- rncols

rse_md_new <- dplyr::select(left_join(rse_md, pdadd, by = 'RNum'), 
                     'SAMPLE_ID', all_of(rncols), everything())
stopifnot(all.equal(rse_md_new$SAMPLE_ID, rownames(rse_md)))
rownames(rse_md_new) <- rse_md_new$SAMPLE_ID
```


```{r}
## and finally we need to assign colData back to the rse_gene
## NOTE: we have to convert the data frame back to Bioconductor's DataFrame type!
colData(rse_gene) <- as(rse_md_new, "DataFrame")
save(rse_gene, file='RTI/rse_gene_RTI_n242.Rdata')
rm(rse_gene)
```


```{r}
load('RTI/rse_exon_Jlab_experiment_n242.Rdata')
rse_md$RNum <- NULL
all.equal(as.data.frame(colData(rse_exon)), rse_md)

colData(rse_exon) <- as(rse_md_new, "DataFrame")
save(rse_exon, file='RTI/rse_exon_RTI_n242.Rdata')
rm(rse_exon)
```

```{r}
load('RTI/rse_tx_Jlab_experiment_n242.Rdata')
stopifnot(all.equal(as.data.frame(colData(rse_tx)), rse_md))

colData(rse_tx) <- as(rse_md_new, "DataFrame")
save(rse_tx, file='RTI/rse_tx_RTI_n242.Rdata')
rm(rse_tx)
```

```{r}
load('RTI/rse_jx_Jlab_experiment_n242.Rdata')
stopifnot(all.equal(as.data.frame(colData(rse_jx)), rse_md))

colData(rse_jx) <- as(rse_md_new, "DataFrame")
save(rse_jx, file='RTI/rse_jx_RTI_n242.Rdata')
rm(rse_jx)
```

#--- What if we want to rearrange the columns? Here it is a simple function that just takes a column name
#--- and the new position it should be in
##arrange data frame vars by position
##'vars' must be a named vector, e.g. c("var.name"=1)
arrange.vars <- function(data, vars){
    ##stop if not a data.frame (but should work for matrices as well)
    stopifnot(is.data.frame(data))
    ##sort out inputs
    data.nms <- names(data)
    var.nr <- length(data.nms)
    var.nms <- names(vars)
    var.pos <- vars
    ##sanity checks
    stopifnot( !any(duplicated(var.nms)), 
               !any(duplicated(var.pos)) )
    stopifnot( is.character(var.nms), 
               is.numeric(var.pos) )
    stopifnot( all(var.nms %in% data.nms) )
    stopifnot( all(var.pos > 0), 
               all(var.pos <= var.nr) )

    ##prepare output
    out.vec <- character(var.nr)
    out.vec[var.pos] <- var.nms
    out.vec[-var.pos] <- data.nms[ !(data.nms %in% var.nms) ]
    stopifnot( length(out.vec)==var.nr )

    ##re-arrange vars by position
    data <- data[ , out.vec]
    return(data)
}
```

```{r}
## now we can rearrange the columns:
rse_mdre <- arrange.vars(rse_md, c("BrNum"=2))
rse_mdre <- arrange.vars(rse_mdre, c("Race"=3))
rse_mdre <- arrange.vars(rse_mdre, c("Age"=4))
rse_mdre <- arrange.vars(rse_mdre, c("Sex"=5))
rse_mdre <- arrange.vars(rse_mdre, c("Dx"=6))
```

